---
title: "Wet SOC Persistence: Data processing"
format: 
    html:
        embed-resources: false
---

```{r}
#| label: setup
library(tidyverse)
library(formatR)
library(openxlsx)
library(here)

knitr::opts_chunk$set(fig.align = "center", fig.show = "hold", dpi = 100,
                      tidy.opts = list(width.cutoff = 60), 
                      tidy = T, 
                      collapse = TRUE, 
                      echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)

rgl::setupKnitr(autoprint = TRUE)
here::here()
```

## Import lab excel datasheets from local drive on computer

```{r}

HLEF_HPU <- openxlsx::read.xlsx("/Users/Anthony/OneDrive - UW/University of Washington/Lab Work/HLEF2022/HLEF2022LAB.xlsx", sheet = "LOCATION") |> 
  dplyr::select(sample_name, HPU) |> 
  dplyr::rename(sample_ID = sample_name)

readr::write_csv(HLEF_HPU, here("data/dataframes/HLEF_HPUonly.csv"))


HLEF_SOC_All <- openxlsx::read.xlsx("/Users/Anthony/OneDrive - UW/University of Washington/Lab Work/HLEF2022/HLEF2022LAB.xlsx", sheet = "HLEF_FullData") |> 
  dplyr::select(c(1:14)) |>
  dplyr::mutate(top = case_when(is.na(depth_cm - lag(depth_cm)) ~ 0,
                                (depth_cm - lag(depth_cm)) < 0 ~ 0, 
                                .default = lag(depth_cm)), 
                bottom = depth_cm, 
                #center = abs(top - (top - bottom)/2),
                thick = case_when(is.na(depth_cm - lag(depth_cm)) ~ bottom,
                                  (bottom-top) < 0 ~ bottom,
                                  #bottom > 100 & top < 100 ~ bottom - (bottom -100)-lag(bottom),
                                  #bottom > 100 & top == 0 ~ bottom - (bottom -100)-0,
                                  #bottom > 100 & top > 100 ~ 0,
                                  #bottom < 100 & bottom - lag(bottom) < 0 ~ bottom,
                                  #bottom < 100 & bottom - lag(bottom) > 0 ~ bottom - lag(bottom),
                                  #bottom == 100 ~ bottom - lag(bottom),
                                  .default = bottom-top),
                SOC_hor_stock = 100*((carbon_perc/100)*BD_g_cm3*thick*(1-(rock_perc/100))))
readr::write_csv(HLEF_SOC_All, here("data/dataframes/HLEF_SOC_AllOrgMin.csv"))

```


### Water Extractable Organic Carbon

```{r}
toc_batch1 <- read.csv("/Users/Anthony/OneDrive - UW/University of Washington/Lab Work/SCGSR/data/TOC/Anthony_Batch1_010224.csv")
toc_batch2 <- read.csv("/Users/Anthony/OneDrive - UW/University of Washington/Lab Work/SCGSR/data/TOC/Anthony_Batch2_010324.csv")
numName <- read.csv(here("data/dataframes/HLEF_TOC_NumName.csv")) |> mutate(Num = as.character(Num))
glimpse(toc_batch1)

toc_batch1_form <- toc_batch1 |> dplyr::select(Name, TIC..mg.l., TOC..Diff....mg.l.) |> 
    dplyr::filter(!str_detect(Name, "Blank|blank|RunIn|standard")) |> 
    na.omit() |> 
    group_by(Name) |> 
    summarise(Num = first(Name), 
              TIC = mean(TIC..mg.l.),
              TOC = mean(TOC..Diff....mg.l.)) |> 
    dplyr::select(-Name)
toc_batch2_form <- toc_batch2 |> dplyr::select(Name, TIC..mg.l., TOC..Diff....mg.l.) |> 
    dplyr::filter(!str_detect(Name, "Blank|blank|RunIn|standard")) |> 
    na.omit() |> 
    group_by(Name) |> 
    summarise(Num = first(Name), 
              TIC = mean(TIC..mg.l.),
              TOC = mean(TOC..Diff....mg.l.)) |> 
    dplyr::select(-Name)

toc_all <- rbind(toc_batch1_form, toc_batch2_form) |>
    left_join(y = numName, by = join_by(Num)) |>
    dplyr::relocate(Name, Num, TIC, TOC) |> 
    filter(Name != "Blank")

readr::write_csv(toc_all, here("data/dataframes/HLEF_TOC_MineralSoils.csv"))

```


### Data filtering for Heavy Fraction Fine (HFF):

- filter for HFF data
- add columns for HFF mgC/gSoil and silt+clay

```{r}
HLEF_Full <- openxlsx::read.xlsx("/Users/Anthony/OneDrive - UW/University of Washington/Lab Work/HLEF2022/HLEF2022LAB.xlsx", sheet = "HLEF_FullData") |>
    mutate(top = case_when(is.na(depth_cm - lag(depth_cm)) ~ 0,
                               .default = lag(depth_cm)), 
            bottom = depth_cm, 
            center = abs(top - (top - bottom)/2),
            thick = case_when(is.na(depth_cm - lag(depth_cm)) ~ bottom,
                              bottom - lag(bottom) < 0 ~ bottom,
                              bottom - lag(bottom) > 0 ~ bottom - lag(bottom)),
            thick100 = case_when(is.na(depth_cm - lag(depth_cm)) ~ bottom,
                                    bottom > 100 & top < 100 ~ bottom - (bottom -100)-lag(bottom),
                                    bottom > 100 & top == 0 ~ bottom - (bottom -100)-0,
                                    bottom > 100 & top > 100 ~ 0,
                                    bottom < 100 & bottom - lag(bottom) < 0 ~ bottom,
                                    bottom < 100 & bottom - lag(bottom) > 0 ~ bottom - lag(bottom),
                                    bottom == 100 ~ bottom - lag(bottom),
                                    .default = 0),
           SOC_hor_stock100 = 100*((carbon_perc/100)*BD_g_cm3*thick100*(1-(rock_perc/100))),
           SOC_hor_stock = 100*((carbon_perc/100)*BD_g_cm3*thick*(1-(rock_perc/100))),
           HFF_SOC_hor_stock_100 =
               100*((HFF_percC/100)*BD_g_cm3*thick100*(1-(rock_perc/100))*((HFF_perc))),
           HFF_SOC_hor_stock =
               100*((HFF_percC/100)*BD_g_cm3*thick*(1-(rock_perc/100))*((HFF_perc)))
           #name_short = paste0(str_sub(sample_ID, start = 1L, end = 3L))
           )

HLEF_loc <- openxlsx::read.xlsx("/Users/Anthony/OneDrive - UW/University of Washington/Lab Work/HLEF2022/HLEF2022LAB.xlsx", sheet = "LOCATION", detectDates = TRUE) |> 
    filter(sample_name != "PRE-2022")

l_n <- lm((nitrogen_perc) ~ carbon_perc + 0, data = HLEF_Full |> filter(nitrogen_perc >0) |> na.omit())
p_n <- predict(l_n, data.frame(carbon_perc = HLEF_Full[HLEF_Full$nitrogen_perc == 0, "carbon_perc"][1]))

HLEF_HFF <- HLEF_Full |> 
    left_join(HLEF_loc |> select(sample_name, HPU),
              by = join_by(sample_ID == sample_name)) |> 
    filter(!is.na(HFF_percC)) |> 
    left_join(y = toc_all |> dplyr::select(Name, TIC, TOC), by = join_by(HLEF_ID == Name)) |> 
    mutate(SOC_mgC_gSoil = (carbon_perc/100)*1000,
           HFF_mgC_gSoil = (HFF_percC/100)*(HFF/df_start_wt)*1000,
           HFF_percC_of_soilC = ((HFF_percC/100)*HFF)/((carbon_perc/100)*df_start_wt),
           HFF_CN = HFF_percC/HFF_percN,
           siltclay = silt+clay,
           depth_cm = center,
           nitrogen_perc = case_when(nitrogen_perc == 0 ~ 0.01,
                              .default = nitrogen_perc),
           bulkCN = carbon_perc/nitrogen_perc) |> 
    dplyr::select(sample_ID, HLEF_ID, HPU, lat, lon, depth_cm, BD_g_cm3, rock_perc, 
                  carbon_perc, TIC, TOC, bulkCN,
                  nitrogen_perc, ph, clay, siltclay, silt, sand, 
                  SOC_mgC_gSoil, SOC_hor_stock, SOC_hor_stock100,
                  HFF_percC, HFF_mgC_gSoil, HFF_CN, HFF_SOC_hor_stock, HFF_SOC_hor_stock_100)



readr::write_csv(HLEF_HFF, here("data/dataframes/HLEF_HFF_SOConly.csv"))
```


```{r}
target_maom14C <- c("HFF_mgC_gSoil", "delta_14C", "delta_14C_adj")

ICP_DC <- openxlsx::read.xlsx("/Users/Anthony/OneDrive - UW/University of Washington/Lab Work/SCGSR/data/ICP_MS/LLNL_HLEF_Extracts_ICP_MS.xlsx", sheet = 3, detectDates = TRUE) |> 
    filter(Extract == "DithCit") |> 
    mutate(Al_mg_gSoil = (Al_ppm*(extract_vol_ml/1000))/(weight_g),
           Fe_mg_gSoil = (Fe_ppm*(extract_vol_ml/1000))/(weight_g),
           Si_mg_gSoil = (Si_ppm*(extract_vol_ml/1000))/(weight_g),
           Ca_mg_gSoil = (Ca_ppm*(extract_vol_ml/1000))/(weight_g),
           Mn_mg_gSoil = (Mn_ppm*(extract_vol_ml/1000))/(weight_g))
ICP_AO <- openxlsx::read.xlsx("/Users/Anthony/OneDrive - UW/University of Washington/Lab Work/SCGSR/data/ICP_MS/LLNL_HLEF_Extracts_ICP_MS.xlsx", sheet = 3, detectDates = TRUE) |> 
    filter(Extract == "AmmOx")  |> 
    mutate(Al_mg_gSoil = (Al_ppm*(extract_vol_ml/1000))/(weight_g),
           Fe_mg_gSoil = (Fe_ppm*(extract_vol_ml/1000))/(weight_g),
           Si_mg_gSoil = (Si_ppm*(extract_vol_ml/1000))/(weight_g),
           Ca_mg_gSoil = (Ca_ppm*(extract_vol_ml/1000))/(weight_g),
           Mn_mg_gSoil = (Mn_ppm*(extract_vol_ml/1000))/(weight_g))

CAMS_14C <- openxlsx::read.xlsx("/Users/Anthony/OneDrive - UW/University of Washington/Lab Work/SCGSR/data/14C/14C_Dataframe.xlsx", sheet = 1) |> 
    dplyr::rename("delta_14C" = "∆14C") |>
    arrange(desc(HLEF_ID))

HLEF_HFF_14C_DithCit_df <- HLEF_HFF |> 
    dplyr::left_join(y = CAMS_14C, by = join_by("HLEF_ID")) |>
    dplyr::left_join(y = ICP_DC, by = join_by("HLEF_ID")) |> 
    #dplyr::left_join(y = ICP_AO, by = join_by("HLEF_ID"), suffix = c("", "_AO")) |> 
    dplyr::filter(!is.na(Fe_ppm)) |> 
    dplyr::select(-ends_with("ppm"), -starts_with("extract"))
HLEF_HFF_14C_AmmOx_df <- HLEF_HFF |> 
    dplyr::left_join(y = CAMS_14C, by = join_by("HLEF_ID")) |>
    #dplyr::left_join(y = ICP_DC, by = join_by("HLEF_ID")) |> 
    dplyr::left_join(y = ICP_AO, by = join_by("HLEF_ID")) |> 
    filter(!is.na(Fe_ppm)) |> 
    dplyr::select(-ends_with("ppm"), -starts_with("extract"))
HLEF_HFF_14C_DithCit_df
HLEF_HFF_14C_AmmOx_df

readr::write_csv(HLEF_HFF_14C_DithCit_df, here("data/dataframes/HLEF_HFF_14C_DithCit_df.csv"))
readr::write_csv(HLEF_HFF_14C_AmmOx_df, here("data/dataframes/HLEF_HFF_14C_AmmOx_df.csv"))


HLEF_HFF_14C_DC_AO <- HLEF_HFF_14C_AmmOx_df |> 
  dplyr::rename_with(~ str_replace(., "_mg_gSoil", "_mg_gSoil_AO"), 
                     ends_with("_mg_gSoil")) |> 
  dplyr::left_join(HLEF_HFF_14C_DithCit_df |> 
                     dplyr::rename_with(~ str_replace(., "_mg_gSoil", "_mg_gSoil_DC"), 
                     ends_with("_mg_gSoil")) |> 
                     dplyr::select(HLEF_ID, ends_with("_mg_gSoil_DC")),
                   by = join_by(HLEF_ID)
                   ) |> 
  dplyr::mutate(Al_AODC_ratio = Al_mg_gSoil_AO/Al_mg_gSoil_DC,
                Fe_AODC_ratio = Fe_mg_gSoil_AO/Fe_mg_gSoil_DC,
                delta_14C_adj = 1000+delta_14C)


readr::write_csv(HLEF_HFF_14C_DC_AO, here("data/dataframes/HLEF_HFF_14C_DC_AO.csv"))

```


```{r}
HLEF_HFF |> group_by(sample_ID) |> summarise(hpu = first(HPU)) |> group_by(hpu) |> summarise(n = n())


```


```{r}
HLEF_SOC_stock <- HLEF_SOC_All |> 
    dplyr::group_by(sample_ID) |> 
    summarise(sample_ID = first(sample_ID),
              SOC_stock = sum(SOC_hor_stock))
readr::write_csv(HLEF_SOC_stock, file = here("data/dataframes/HLEF_totSOCstock.csv"))

HLEF_HFF_stock <- HLEF_HFF |>
    dplyr::group_by(sample_ID) |> 
    summarise(sample_ID = first(sample_ID),
              HFF_Stock = sum(HFF_SOC_hor_stock))
# readr::write_csv(HLEF_HFF_stock, file = "WetSOC_Persistence/data/dataframes/HLEF_HFFstock.csv")

HLEF_stock_cmb <- HLEF_SOC_stock |> 
    dplyr::left_join(HLEF_HFF_stock, by = join_by(sample_ID)) |> 
    dplyr::left_join(HLEF_HPU, by = join_by(sample_ID)) |> 
    dplyr::mutate(HFF_Stock = case_when(is.na(HFF_Stock) ~ 0,
                                 .default = HFF_Stock),
                  HFF_SOC_ratio = HFF_Stock/SOC_stock)

# readr::write_csv(HLEF_stock_cmb, file = "WetSOC_Persistence/data/dataframes/HLEF_HFFtotSOCstock_combined.csv")

```


## End summary

Should have: 
1. A dataframe with `sample_ID` and `HPU` classes for all 84 samples (Organic and Mineral)
2. A dataframe with basic lab data but no SCGSR data 
3. A dataframe of TIC and TOC values for 50 mineral horizons
4. A dataframe with heavy-fine fraction (HFF) for 50 mineral horizons
5. A dataframe of Dithionate-citrate extract measurements for 50 mineral horizons
6. A dataframe of Ammonium-oxalate extract measurements for 50 mineral horizons
7. A dataframe of both DC and AO measurements with the 14C measurements combined 
  - This was not combined earlier due to missing analysis
  
Analysis can be run on #7 dataframe

```{r}
read.csv(here("data/dataframes/HLEF_HPUonly.csv")) |> glimpse() #1
read.csv(here("data/dataframes/HLEF_SOC_AllOrgMin.csv")) |> glimpse() #2
read.csv(here("data/dataframes/HLEF_TOC_MineralSoils.csv")) |> glimpse() #3
read.csv(here("data/dataframes/HLEF_HFF_SOConly.csv")) |> glimpse() #4
read.csv(here("data/dataframes/HLEF_HFF_14C_DithCit_df.csv")) |> glimpse() #5
read.csv(here("data/dataframes/HLEF_HFF_14C_AmmOx_df.csv")) |> glimpse() #6
read.csv( here("data/dataframes/HLEF_HFF_14C_DC_AO.csv")) |> glimpse() #7 
```

2. 

```{r}

```

