---
title: "Wetland SOC & MAOM Analysis"
format: html
---

```{r}
#| label: setup

library(tidyverse)
library(lme4)
library(car)
library(MASS)
library(lmerTest)
library(emmeans)
library(ppcor)
library(ggcorrplot)
library(RColorBrewer)
library(webshot)
library(kableExtra)
library(formatR)
library(openxlsx)
library(here)

knitr::opts_chunk$set(fig.align = "center", fig.show = "hold", dpi = 100,
                      tidy.opts = list(width.cutoff = 60), 
                      tidy = T, 
                      collapse = TRUE, 
                      echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)
knitr::opts_knit$set(root.dir = '/Users/Anthony/OneDrive - UW/University of Washington/Data and Modeling/')

#knitr::knit_hooks$set(webgl = hook_webgl)
rgl::setupKnitr(autoprint = TRUE)

r.sq <- function(y,y.fitted){
    res <- y-y.fitted
    1-sum(res^2)/sum((y-mean(y))^2)
}

check_assump_func <- function(model) {
  plot(model)
  plot(model, resid(.) ~ fitted(.)|sample_ID, abline = 0, ylab = "Model Residuals", xlab = "Fitted Model")
  
  qqnorm(resid(model))
  qqline(resid(model))
  
  lev <- hat(model.matrix(model))
  plot(resid(model) ~ lev, xlab = "Leverage", ylab = "Model Residuals")
  
  cd <- cooks.distance(model)
  plot(x = lev, y = cd, xlab = "Leverage", ylab = "Cooks Distance", ylim = c(0,1))
  # legend("topleft",
  #        legend = "Cooks Distance", 
  #        text.col = "blue",
  #        pch = 1,
  #        col = "blue")
  
  hist(as.vector(unlist(ranef(model)$sample_ID)), main = "", xlab = "Random Effect Conditional Means")
}

```


data import 
```{r}
HLEF_HPU <- read.csv(here("data/dataframes/HLEF_HPUonly.csv")) #|> glimpse() #1
HLEF_SOC_All <- read.csv(here("data/dataframes/HLEF_SOC_AllOrgMin.csv")) #|> glimpse() #2
HLEF_TOC <- read.csv(here("data/dataframes/HLEF_TOC_MineralSoils.csv")) #|> glimpse() #3
HLEF_HFF <- read.csv(here("data/dataframes/HLEF_HFF_SOConly.csv")) #|> glimpse() #4
HLEF_HFF_14C_DithCit_df <- read.csv(here("data/dataframes/HLEF_HFF_14C_DithCit_df.csv")) #|> glimpse() #5
HLEF_HFF_14C_AmmOx_df <- read.csv(here("data/dataframes/HLEF_HFF_14C_AmmOx_df.csv")) # |> glimpse() #6
HLEF_HFF_14C_DC_AO_trans <- read.csv( here("data/dataframes/HLEF_HFF_14C_DC_AO.csv")) #|> glimpse() #7 
```


MAOM-C content(mg/g) between HPUs or depth?

```{r}
maoc_lm <- glmer((HFF_mgC_gSoil) ~ (HPU) + (1|sample_ID), 
                 data = HLEF_HFF, family = Gamma(link = "log"))
summary(maoc_lm)
car::Anova(maoc_lm, type = 2)

#TukeyHSD(maoc_lm, "HPU")

emm_maoc <- emmeans(maoc_lm, specs = pairwise ~ HPU, adjust = "tukey")
emm_maoc

```


```{r}

maoc_depthHPU_lm <- glmer((HFF_mgC_gSoil) ~ HPU*depth_cm +(1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
summary(maoc_depthHPU_lm)
Anova(maoc_depthHPU_lm)
emm_maoc_depthHPU_lm <- emtrends(maoc_depthHPU_lm, specs = pairwise ~ HPU, var = "depth_cm",
                               regrid = "response", adjust = "tukey")
emm_maoc_depthHPU_lm
emmip(maoc_depthHPU_lm, HPU~depth_cm, cov.reduce = range)

```

MAOM-C content with SOC  content

```{r}
maoc_soc <- glmer((HFF_mgC_gSoil) ~ SOC_mgC_gSoil +(1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
summary(maoc_soc)

HLEF_HFF_maom_soc <- HLEF_HFF |> 
  dplyr::mutate(MAOM_SOC_Diff = SOC_mgC_gSoil - HFF_mgC_gSoil)



ggplot(data = HLEF_HFF_maom_soc) +
  geom_point(aes(y=HFF_mgC_gSoil, x = SOC_mgC_gSoil, fill = HPU), shape = 21) +
  geom_smooth(aes(y=HFF_mgC_gSoil, x = SOC_mgC_gSoil, colour = HPU), method = "lm")
```


MAOM-C content(mg/g) correlations with Bulk Soil properties

```{r}
HLEF_cor <- round(cor(HLEF_HFF_14C_DC_AO_trans |> dplyr::select(HFF_mgC_gSoil, depth_cm, carbon_perc,
                                                TOC, nitrogen_perc, clay, siltclay, ph) ), 1)
HLEF_cor_pmat <- cor_pmat(HLEF_cor)

ggcorrplot(HLEF_cor, type = "lower", p.mat = HLEF_cor_pmat, 
           sig.level = 0.05, lab = TRUE)

```

### MAOM-C Bulk Soil modeling

```{r}
lmC_MAOM <- glmer((HFF_mgC_gSoil) ~ carbon_perc + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
summary(lmC_MAOM)
Anova(lmC_MAOM, type = 3)

# lmTOC_MAOM <- glmer((HFF_mgC_gSoil) ~ TOC + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
# summary(lmTOC_MAOM)
# Anova(lmTOC_MAOM, type = 3)

# lmN_MAOM <- glmer((HFF_mgC_gSoil) ~ bulkCN + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
# Anova(lmN_MAOM, type = 3)

lmClay_MAOM <- glmer((HFF_mgC_gSoil) ~ clay + (1|sample_ID), 
                     data = HLEF_HFF_14C_DC_AO_trans, 
                     family = Gamma(link = "log"),
                     control = glmerControl(optimizer = "nloptwrap"))
summary(lmClay_MAOM)
Anova(lmClay_MAOM, type = 3)

# lmSiltClay_MAOM <- glmer((HFF_mgC_gSoil) ~ siltclay + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
# Anova(lmSiltClay_MAOM, type = 3)

lmpH_MAOM <- glmer((HFF_mgC_gSoil) ~ ph + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, 
                   family = Gamma(link = "log"))
summary(lmpH_MAOM)
Anova(lmpH_MAOM, type = 3)

lmbulk_MAOM_HPU <- glmer((HFF_mgC_gSoil) ~ carbon_perc + clay + ph + HPU + (1|sample_ID), 
                     data = HLEF_HFF_14C_DC_AO_trans, 
                     family = Gamma(link = "log"),
                     control = glmerControl(optimizer = "nloptwrap"))
lmbulk_MAOM <- glmer((HFF_mgC_gSoil) ~ carbon_perc + clay + ph + (1|sample_ID), 
                     data = HLEF_HFF_14C_DC_AO_trans, 
                     family = Gamma(link = "log"),
                     control = glmerControl(optimizer = "nloptwrap"))
summary(lmbulk_MAOM)
summary(lmbulk_MAOM_HPU)
Anova(lmbulk_MAOM, type = 3)
Anova(lmbulk_MAOM_HPU, type = 3)


```



### MAOM-C modeling to find important variables

```{r}
lmAll_MAOM_HPU <- glmer((HFF_mgC_gSoil) ~ carbon_perc + clay + ph + HPU + 
                          Fe_mg_gSoil_AO + Fe_mg_gSoil_DC + Al_mg_gSoil_AO +
                          (1|sample_ID), 
                     data = HLEF_HFF_14C_DC_AO_trans, 
                     family = Gamma(link = "log"),
                     control = glmerControl(optimizer = "Nelder_Mead"))
summary(lmAll_MAOM_HPU)
Anova(lmAll_MAOM_HPU, type = 3)
```

Parsimonious model
```{r}
lmAll_MAOM <- glmer((HFF_mgC_gSoil) ~ carbon_perc + ph + 
                          Fe_mg_gSoil_DC + 
                          (1|sample_ID), 
                     data = HLEF_HFF_14C_DC_AO_trans, 
                     family = Gamma(link = "log"),
                     control = glmerControl(optimizer = "Nelder_Mead"))
summary(lmAll_MAOM)
Anova(lmAll_MAOM, type = 3)
AIC(lmAll_MAOM)

saveRDS(lmAll_MAOM, file = "WetSOC_Persistence/analysis/models/MAOM_All_selection.rds")

```


Linear models for examining relationships with MAOM-C content
```{r}
print("***************Fe***************")
lmFeDC_MAOM <- glmer((HFF_mgC_gSoil) ~ Fe_mg_gSoil_DC + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
check_assump_func(lmFeDC_MAOM)
Anova(lmFeDC_MAOM)
lmFeAO_MAOM <- glmer((HFF_mgC_gSoil) ~ Fe_mg_gSoil_AO + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
check_assump_func(lmFeAO_MAOM)
Anova(lmFeAO_MAOM)
lmFeAODC_MAOM <- glmer((HFF_mgC_gSoil) ~ Fe_AODC_ratio + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
check_assump_func(lmFeAODC_MAOM)
Anova(lmFeAODC_MAOM)

print("***************Al***************")
lmAlDC_MAOM <- glmer((HFF_mgC_gSoil) ~ Al_mg_gSoil_DC + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
check_assump_func(lmAlDC_MAOM)
Anova(lmAlDC_MAOM)
lmAlAO_MAOM <- glmer((HFF_mgC_gSoil) ~ Al_mg_gSoil_AO + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
check_assump_func(lmAlAO_MAOM)
Anova(lmAlAO_MAOM)
lmAlAODC_MAOM <- glmer((HFF_mgC_gSoil) ~ Al_AODC_ratio + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
check_assump_func(lmAlAODC_MAOM)
Anova(lmAlAODC_MAOM)
# 
# lmCaDC_MAOM <- glmer((HFF_mgC_gSoil) ~ HPU*Ca_mg_gSoil + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
# Anova(lmCaDC_MAOM)
# 
# lmMnDC_MAOM <- glmer((HFF_mgC_gSoil) ~ HPU*Mn_mg_gSoil + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
# Anova(lmMnDC_MAOM)
# 
# lmSiDC_MAOM <- glmer((HFF_mgC_gSoil) ~ HPU*Si_mg_gSoil + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = Gamma(link = "log"))
# Anova(lmSiDC_MAOM)


```

### Partial correlations between bulk soil properties and MAOM-C mg g-1
- Need to log transform still to match models

```{r}
HLEF_HFF_ppcor <- HLEF_HFF_14C_DC_AO_trans |> 
  dplyr::mutate(poorAll = Fe_mg_gSoil_AO + Al_mg_gSoil_DC) |> 
  dplyr::select(HFF_mgC_gSoil, delta_14C, depth_cm, carbon_perc, clay, ph,
                Fe_mg_gSoil_DC, Al_mg_gSoil_DC, Fe_mg_gSoil_AO, Al_mg_gSoil_AO)

pcor(HLEF_HFF_ppcor)$estimate |> 
  ggcorrplot::ggcorrplot(lab = TRUE, lab_size = 2, 
                         p.mat = pcor(HLEF_HFF_ppcor)$p.value, type = "upper")
```


```{r}
cor.test((HLEF_HFF_ppcor$HFF_mgC_gSoil), HLEF_HFF_ppcor$carbon_perc)
cor.test((HLEF_HFF_ppcor$HFF_mgC_gSoil), HLEF_HFF_ppcor$clay)
cor.test((HLEF_HFF_ppcor$HFF_mgC_gSoil), HLEF_HFF_ppcor$ph)

pcor.test((HLEF_HFF_ppcor$depth_cm), 
          y = HLEF_HFF_ppcor$ph, 
          z = HLEF_HFF_ppcor$carbon_perc)
pcor.test((HLEF_HFF_ppcor$HFF_mgC_gSoil), 
          y = HLEF_HFF_ppcor$carbon_perc, 
          z = HLEF_HFF_ppcor$depth_cm)
pcor.test((HLEF_HFF_ppcor$HFF_mgC_gSoil),
          y = HLEF_HFF_ppcor$clay, 
          z = HLEF_HFF_ppcor$depth_cm)
pcor.test((HLEF_HFF_ppcor$HFF_mgC_gSoil), 
          y = HLEF_HFF_ppcor$ph, 
          z = HLEF_HFF_ppcor$depth_cm)

pcor.test((HLEF_HFF_ppcor$HFF_mgC_gSoil), 
          y = HLEF_HFF_ppcor$Fe_mg_gSoil_DC, 
          z = HLEF_HFF_ppcor$carbon_perc)
pcor.test((HLEF_HFF_ppcor$HFF_mgC_gSoil), 
          y = HLEF_HFF_ppcor$Fe_mg_gSoil_AO, 
          z = HLEF_HFF_ppcor$ph)
pcor.test((HLEF_HFF_ppcor$HFF_mgC_gSoil), 
          y = HLEF_HFF_ppcor$Al_mg_gSoil_DC, 
          z = HLEF_HFF_ppcor$ph)
pcor.test((HLEF_HFF_ppcor$HFF_mgC_gSoil), 
          y = HLEF_HFF_ppcor$Al_mg_gSoil_AO, 
          z = HLEF_HFF_ppcor$ph)


pcor.test((HLEF_HFF_ppcor$HFF_mgC_gSoil), y = HLEF_HFF_ppcor$Fe_mg_gSoil_DC, 
          z = HLEF_HFF_ppcor[, c("ph", "clay")])



```

```{r}
calculate_multiple_correlations <- function(df, target, predictors, control_vars = NULL) {
  # Input validation
  if(!is.data.frame(df)) {
    stop("Input must be a dataframe")
  }
  
  if(!(target %in% names(df))) {
    stop(paste("Target variable", target, "not found in dataframe"))
  }
  
  if(!all(predictors %in% names(df))) {
    missing_predictors <- predictors[!(predictors %in% names(df))]
    stop(paste("Predictor variable(s) not found in dataframe:", 
               paste(missing_predictors, collapse = ", ")))
  }
  
  if(!is.null(control_vars)) {
    missing_controls <- control_vars[!(control_vars %in% names(df))]
    if(length(missing_controls) > 0) {
      stop(paste("Control variable(s) not found in dataframe:", 
                 paste(missing_controls, collapse = ", ")))
    }
  }
  
  # Check for ppcor package
  if(!is.null(control_vars) && length(control_vars) > 0) {
    if(!requireNamespace("ppcor", quietly = TRUE)) {
      stop("Package 'ppcor' is needed for partial correlations. Please install it.")
    }
  }
  
  # Initialize empty result dataframe
  results <- data.frame(
    predictor = character(),
    zero_order_r = numeric(),
    zero_order_pvalue = numeric(),
    control_variables = character(),
    partial_r = numeric(),
    partial_pvalue = numeric(),
    stringsAsFactors = FALSE
  )
  
  # Calculate correlations for each predictor
  for(pred in predictors) {
    # Calculate zero-order correlation
    zero_order_model <- cor.test(df[[target]], df[[pred]])
    zero_order_r <- zero_order_model$estimate
    zero_order_pvalue <- zero_order_model$p.value
    
    # Calculate partial correlation (if control variables are provided)
    if(is.null(control_vars) || length(control_vars) == 0) {
      partial_r <- NA
      partial_pvalue <- NA
      control_vars_str <- NA
    } else {
      # Calculate partial correlation
      partial_result <- ppcor::pcor.test(
        x = df[[pred]],
        y = df[[target]],
        z = df[, control_vars, drop = FALSE]
      )
      
      partial_r <- partial_result$estimate
      partial_pvalue <- partial_result$p.value
      control_vars_str <- paste(control_vars, collapse = ", ")
    }
    
    # Create result row for this predictor
    result_row <- data.frame(
      predictor = pred,
      zero_order_r = zero_order_r,
      zero_order_pvalue = zero_order_pvalue,
      control_variables = ifelse(is.na(control_vars_str), "None", control_vars_str),
      partial_r = partial_r,
      partial_pvalue = partial_pvalue,
      stringsAsFactors = FALSE
    )
    
    # Append to results dataframe
    results <- rbind(results, result_row)
  }
  
  return(results)
}
```


```{r}
bulk_extract_tot_cor <- calculate_multiple_correlations(df = HLEF_HFF_ppcor, 
                       target = "HFF_mgC_gSoil", 
                       predictor = c("depth_cm", "carbon_perc", "ph", "clay"), 
                       control_vars = c("Fe_mg_gSoil_DC", 
                                        "Al_mg_gSoil_AO"))
bulk_extract_poor_cor <- calculate_multiple_correlations(df = HLEF_HFF_ppcor, 
                       target = "HFF_mgC_gSoil", 
                       predictor = c("depth_cm", "carbon_perc", "ph", "clay"), 
                       control_vars = c("Fe_mg_gSoil_AO", 
                                        "Al_mg_gSoil_DC"))
extract_tot_bulk_cor <- calculate_multiple_correlations(df = HLEF_HFF_ppcor, 
                       target = "HFF_mgC_gSoil", 
                       predictor = c("Fe_mg_gSoil_DC", 
                                        "Al_mg_gSoil_AO"), 
                       control_vars = c("depth_cm", "carbon_perc", "ph", "clay"))
extract_poor_bulk_cor <- calculate_multiple_correlations(df = HLEF_HFF_ppcor, 
                       target = "HFF_mgC_gSoil", 
                       predictor = c("Fe_mg_gSoil_AO", 
                                        "Al_mg_gSoil_DC"), 
                       control_vars = c("depth_cm", "carbon_perc", "ph", "clay"))


readr::write_csv(bulk_extract_tot_cor, "WetSOC_Persistence/data/dataframes/bulk_extract_total_cor_pcor_comparison.csv")
readr::write_csv(bulk_extract_poor_cor, "WetSOC_Persistence/data/dataframes/bulk_extract_poorCry_cor_pcor_comparison.csv")
readr::write_csv(extract_tot_bulk_cor, "WetSOC_Persistence/data/dataframes/extract_total_bulk_cor_pcor_comparison.csv")
readr::write_csv(extract_poor_bulk_cor, "WetSOC_Persistence/data/dataframes/extract_poorCry_bulk_cor_pcor_comparison.csv")

```


Is there a difference in the ∆14C between HPUs?

```{r}

maoc_14C_lm <- lmer((delta_14C) ~ HPU + (1|sample_ID),
                     data = HLEF_HFF_14C_DC_AO_trans)
check_assump_func(maoc_14C_lm)
summary(maoc_14C_lm)
Anova(maoc_14C_lm)
emmeans(maoc_14C_lm, pairwise ~ HPU)




```
```{r}
maoc_14C_depth_lm <- lmer((delta_14C) ~ HPU*depth_cm + (1|sample_ID),
                     data = HLEF_HFF_14C_DC_AO_trans)
check_assump_func(maoc_14C_depth_lm)
summary(maoc_14C_depth_lm)
Anova(maoc_14C_depth_lm)
emmeans(maoc_14C_depth_lm, pairwise ~ HPU)
```



14C Correlations 

```{r}
lmC_MAOM_14C <- lmer((delta_14C_adj) ~ carbon_perc + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans)
summary(lmC_MAOM_14C)
check_assump_func(lmC_MAOM_14C)
Anova(lmC_MAOM_14C)

# lmCN_MAOM_14C <- glmer((delta_14C_adj) ~ bulkCN + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = gaussian(link = "identity"))
# Anova(lmCN_MAOM_14C)

lmClay_MAOM_14C <- lmer((delta_14C_adj) ~ clay + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans)
check_assump_func(lmClay_MAOM_14C)
Anova(lmClay_MAOM_14C)

# lmSiltClay_MAOM_14C <- glmer((delta_14C_adj) ~ siltclay + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = gaussian(link = "identity"))
# Anova(lmSiltClay_MAOM_14C)

lmpH_MAOM_14C <- lmer((delta_14C_adj) ~ ph + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans)
summary(lmpH_MAOM_14C)
check_assump_func(lmpH_MAOM_14C)
Anova(lmpH_MAOM_14C)

lmDepth_MAOM_14C <- lmer((delta_14C_adj) ~ depth_cm + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans)
summary(lmDepth_MAOM_14C)
check_assump_func(lmDepth_MAOM_14C)
Anova(lmDepth_MAOM_14C)

```

```{r}
print("***************Fe***************")
lmFeDC_MAOM_14C <- lmer((delta_14C_adj) ~ Fe_mg_gSoil_DC + (1|sample_ID), 
                         data = HLEF_HFF_14C_DC_AO_trans)
check_assump_func(lmFeDC_MAOM_14C)
summary(lmFeDC_MAOM_14C)
Anova(lmFeDC_MAOM_14C)
lmFeAO_MAOM_14C <- lmer((delta_14C_adj) ~ Fe_mg_gSoil_AO + (1|sample_ID), 
                         data = HLEF_HFF_14C_DC_AO_trans)
check_assump_func(lmFeAO_MAOM_14C)
summary(lmFeAO_MAOM_14C)
Anova(lmFeAO_MAOM_14C)
lmFeAODC_MAOM_14C <- lmer((delta_14C_adj) ~ Fe_AODC_ratio + (1|sample_ID), 
                           data = HLEF_HFF_14C_DC_AO_trans)
check_assump_func(lmFeAODC_MAOM_14C)
summary(lmFeAODC_MAOM_14C)
Anova(lmFeAODC_MAOM_14C)

print("***************Al***************")
lmAlDC_MAOM_14C <- lmer((delta_14C_adj) ~ Al_mg_gSoil_DC + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans)
check_assump_func(lmAlDC_MAOM_14C)
summary(lmAlDC_MAOM_14C)
Anova(lmAlDC_MAOM_14C)
lmAlAO_MAOM_14C <- lmer((delta_14C_adj) ~ Al_mg_gSoil_AO + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans)
check_assump_func(lmAlAO_MAOM_14C)
summary(lmAlAO_MAOM_14C)
Anova(lmAlAO_MAOM_14C)
lmAlAODC_MAOM_14C <- lmer((delta_14C_adj) ~ Al_AODC_ratio + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans)
check_assump_func(lmAlAODC_MAOM_14C)
summary(lmAlAODC_MAOM_14C)
Anova(lmAlAODC_MAOM_14C)

# lmCaDC_MAOM_14C <- lmer((delta_14C_adj) ~ Ca_mg_gSoil + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans)
# Anova(lmCaDC_MAOM_14C)
# 
# lmMnDC_MAOM_14C <- glmer((delta_14C_adj) ~ Mn_mg_gSoil + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = gaussian(link = "identity"))
# Anova(lmMnDC_MAOM_14C)
# 
# lmSiDC_MAOM_14C <- glmer((delta_14C_adj) ~ Si_mg_gSoil + (1|sample_ID), data = HLEF_HFF_14C_DC_AO_trans, family = gaussian(link = "identity"))
# Anova(lmSiDC_MAOM_14C)


```



### Partial Correlation 14C

```{r}
bulk_extract_tot_cor14C <- calculate_multiple_correlations(df = HLEF_HFF_ppcor, 
                       target = "delta_14C", 
                       predictor = c("depth_cm", "carbon_perc", "ph", "clay"), 
                       control_vars = c("Fe_mg_gSoil_DC", 
                                        "Al_mg_gSoil_AO"))
bulk_extract_poor_cor14C <- calculate_multiple_correlations(df = HLEF_HFF_ppcor, 
                       target = "delta_14C", 
                       predictor = c("depth_cm", "carbon_perc", "ph", "clay"), 
                       control_vars = c("Fe_mg_gSoil_AO", 
                                        "Al_mg_gSoil_DC"))
extract_tot_bulk_cor14C <- calculate_multiple_correlations(df = HLEF_HFF_ppcor, 
                       target = "delta_14C", 
                       predictor = c("Fe_mg_gSoil_DC", 
                                        "Al_mg_gSoil_AO"), 
                       control_vars = c("depth_cm", "carbon_perc", "ph", "clay"))
extract_poor_bulk_cor14C <- calculate_multiple_correlations(df = HLEF_HFF_ppcor, 
                       target = "delta_14C", 
                       predictor = c("Fe_mg_gSoil_AO", 
                                        "Al_mg_gSoil_DC"), 
                       control_vars = c("depth_cm", "carbon_perc", "ph", "clay"))


readr::write_csv(bulk_extract_tot_cor14C, "WetSOC_Persistence/data/dataframes/bulk_extract_total_cor14C_pcor_comparison.csv")
readr::write_csv(bulk_extract_poor_cor14C, "WetSOC_Persistence/data/dataframes/bulk_extract_poorCry_cor14C_pcor_comparison.csv")
readr::write_csv(extract_tot_bulk_cor14C, "WetSOC_Persistence/data/dataframes/extract_total_bulk_cor14C_pcor_comparison.csv")
readr::write_csv(extract_poor_bulk_cor14C, "WetSOC_Persistence/data/dataframes/extract_poorCry_bulk_cor14C_pcor_comparison.csv")


```

