---
title: "Wetland SOC & MAOM Analysis"
format: html
---

```{r}
#| label: setup

library(dplyr)
library(tidyr)
library(lme4)
library(car)
library(MASS)
library(lmerTest)
library(emmeans)
library(ppcor)
library(ggcorrplot)
library(RColorBrewer)
library(webshot)
library(kableExtra)
library(formatR)
library(openxlsx)
library(broom.mixed)
library(gt)
library(here)

knitr::opts_chunk$set(fig.align = "center", fig.show = "hold", dpi = 100,
                      tidy.opts = list(width.cutoff = 60), 
                      tidy = T, 
                      collapse = TRUE, 
                      echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)

#knitr::knit_hooks$set(webgl = hook_webgl)
rgl::setupKnitr(autoprint = TRUE)
source(here("analysis/functions.R"))
```


data import 
```{r}
HLEF_HPU <- read.csv(here("data/dataframes/HLEF_HPUonly.csv")) #|> glimpse() #1
HLEF_SOC_All <- read.csv(here("data/dataframes/HLEF_SOC_AllOrgMin.csv")) #|> glimpse() #2
HLEF_TOC <- read.csv(here("data/dataframes/HLEF_TOC_MineralSoils.csv")) #|> glimpse() #3
HLEF_HFF <- read.csv(here("data/dataframes/HLEF_HFF_SOConly.csv")) #|> glimpse() #4
HLEF_HFF_14C_DithCit_df <- read.csv(here("data/dataframes/HLEF_HFF_14C_DithCit_df.csv")) #|> glimpse() #5
HLEF_HFF_14C_AmmOx_df <- read.csv(here("data/dataframes/HLEF_HFF_14C_AmmOx_df.csv")) # |> glimpse() #6
HLEF_HFF_14C_DC_AO <- read.csv( here("data/dataframes/HLEF_HFF_14C_DC_AO.csv")) #|> glimpse() #7 
HLEF_HFF_14C_DC_AO_scale <- read.csv(here("data/dataframes/HLEF_HFF_14C_DC_AO_scale.csv")) #|> glimpse() #8
```


### Bulk soil measurements 

```{r}
HLEF_HFF |> group_by(HPU) |> 
    mutate(HPU = as.factor(HPU)) |> 
    summarise(meanC = mean(SOC_mgC_gSoil),
              meanTOC = mean(TOC), 
              meanN = mean(nitrogen_perc),
              meanBCN = mean(bulkCN),
              meanpH = mean(ph),
              meanClay = mean(clay),
              meanSiltClay = mean(siltclay),
              meanDepthOrg = mean(depth_from_org),
              sdC = sd(carbon_perc),
              sdTOC = sd(TOC),
              sdN = sd(nitrogen_perc),
              sdBCN = sd(bulkCN),
              sdpH = sd(ph),
              sdClay = sd(clay),
              sdSiltClay = sd(siltclay),
              n = n()
              ) |> 
    mutate(seC = sdC/sqrt(n),
           seTOC = sdTOC/sqrt(n),
              seN = sdN/sqrt(n),
              seBCN = sdBCN/sqrt(n),
              sepH = sdpH/sqrt(n),
              seClay = sdClay/sqrt(n),
              seSiltClay = sdSiltClay/sqrt(n)
           ) |> 
    dplyr::select(-starts_with("sd")) |> 
    ggplot(aes(x = HPU, y = meanDepthOrg)) +
    geom_col()
```

### Quick summary of predictor models by HPU

```{r}
summarize_glmm(HLEF_HFF, response = "SOC_mgC_gSoil", predictor = "HPU")
summarize_glmm(HLEF_HFF, response = "TOC", predictor = "HPU")
summarize_glmm(HLEF_HFF, response = "nitrogen_perc", predictor = "HPU")
summarize_glmm(HLEF_HFF, response = "bulkCN", predictor = "HPU")
summarize_glmm(HLEF_HFF, response = "ph", predictor = "HPU")
summarize_glmm(HLEF_HFF, response = "clay", predictor = "HPU")
summarize_glmm(HLEF_HFF, response = "siltclay", predictor = "HPU")
summarize_glmm(HLEF_HFF, response = "depth_from_org", predictor = "HPU")
```

### HPU vars with significant differencs 

```{r}
SOC_mod <- glmer(SOC_mgC_gSoil ~ HPU + (1|sample_ID), data = HLEF_HFF, family = Gamma(link = "log"))
check_assump_func(SOC_mod)
summary(SOC_mod)
```



### Extractable Metals 

```{r}
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "Fe_mg_gSoil_DC", predictor = "HPU")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "Fe_mg_gSoil_AO", predictor = "HPU")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "Fe_AODC_ratio", predictor = "HPU")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "Al_mg_gSoil_DC", predictor = "HPU")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "Al_mg_gSoil_AO", predictor = "HPU")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "Al_AODC_ratio", predictor = "HPU")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "Ca_mg_gSoil_AO", predictor = "HPU")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "Ca_mg_gSoil_DC", predictor = "HPU")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "Mn_mg_gSoil_AO", predictor = "HPU")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "Mn_mg_gSoil_DC", predictor = "HPU")
```

```{r}
HPU_univariate <- HLEF_HFF_14C_DC_AO |> 
  dplyr::select(HFF_mgC_gSoil, SOC_mgC_gSoil, nitrogen_perc, ph, depth_cm, depth_from_org,
                HPU, Fe_mg_gSoil_AO, Fe_mg_gSoil_DC, Fe_AODC_ratio, Al_mg_gSoil_DC, 
                Al_AODC_ratio, Ca_mg_gSoil_DC, Mn_mg_gSoil_DC)
readr::write_csv(HPU_univariate, here("Data/dataframes/HPU_univariate.csv"))
```


### MAOM-C content(mg/g) between HPUs or depth?

```{r}
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "HPU")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "depth_cm")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "depth_cm*HPU")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "depth_from_org")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "depth_from_org*HPU")
```


```{r}
maoc_depthHPU_lm <- glmer((HFF_mgC_gSoil) ~ HPU*depth_cm +(1|sample_ID), data = HLEF_HFF_14C_DC_AO, family = Gamma(link = "log"))
emmip(maoc_depthHPU_lm, HPU~depth_cm, cov.reduce = range)

maoc_depthorgHPU_lm <- glmer((HFF_mgC_gSoil) ~ HPU*depth_from_org +(1|sample_ID), data = HLEF_HFF_14C_DC_AO, family = Gamma(link = "log"))
emmip(maoc_depthorgHPU_lm, HPU~depth_from_org, cov.reduce = range)
```


MAOM-C content with SOC  content

```{r}
maoc_soc <- glmer((HFF_mgC_gSoil) ~ SOC_mgC_gSoil +(1|sample_ID), data = HLEF_HFF_14C_DC_AO, family = Gamma(link = "log"))
summary(maoc_soc)

HLEF_HFF_maom_soc <- HLEF_HFF |> 
  dplyr::mutate(MAOM_SOC_Diff = SOC_mgC_gSoil - HFF_mgC_gSoil)

ggplot(data = HLEF_HFF_maom_soc) +
  geom_point(aes(y=HFF_mgC_gSoil, x = SOC_mgC_gSoil, fill = HPU), shape = 21) +
  geom_smooth(aes(y=HFF_mgC_gSoil, x = SOC_mgC_gSoil, colour = HPU), method = "lm")
```


MAOM-C content(mg/g) correlations with Bulk Soil properties

```{r}
HLEF_cor <- round(cor(HLEF_HFF_14C_DC_AO |> 
                        dplyr::select(HFF_mgC_gSoil, depth_cm, depth_from_org, SOC_mgC_gSoil,
                                      TOC, nitrogen_perc, clay, siltclay, ph, 
                                      Al_mg_gSoil_AO, Fe_mg_gSoil_AO, Ca_mg_gSoil_AO,
                                      Al_mg_gSoil_DC, Fe_mg_gSoil_DC, Ca_mg_gSoil_DC) ), 1)
HLEF_cor_pmat <- cor_pmat(HLEF_cor)

corrplot::corrplot(HLEF_cor, method = "pie", type = "upper")

```

### MAOM-C Bulk Soil modeling

```{r}
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "SOC_mgC_gSoil")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "TOC")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "bulkCN")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "clay")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "siltclay")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "ph")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "clay*ph")
```


```{r}
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "Al_mg_gSoil_AO")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "Fe_mg_gSoil_AO")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "Ca_mg_gSoil_AO")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "Al_mg_gSoil_DC")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "Fe_mg_gSoil_DC")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "HFF_mgC_gSoil", predictor = "Ca_mg_gSoil_DC")
```


Provide a dataframe that contains only the response and significant univariate predictor for the FiguresTables.qmd
```{r}
MAOM_univariate <- HLEF_HFF_14C_DC_AO |> 
  dplyr::select(HFF_mgC_gSoil, SOC_mgC_gSoil, TOC, clay, siltclay, ph, depth_cm, depth_from_org,
                HPU,Fe_mg_gSoil_AO, Fe_mg_gSoil_DC, Al_mg_gSoil_AO)
readr::write_csv(MAOM_univariate, here("Data/dataframes/MAOM_univariate.csv"))
```


### ∆14C with bulk soil measurements


```{r}
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "SOC_mgC_gSoil")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "TOC")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "bulkCN")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "clay")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "siltclay")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "depth_cm")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "depth_from_org")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "ph")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "clay*ph")
```

```{r}
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "Al_mg_gSoil_AO")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "Fe_mg_gSoil_AO")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "Ca_mg_gSoil_AO")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "Al_mg_gSoil_DC")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "Fe_mg_gSoil_DC")
summarize_glmm(HLEF_HFF_14C_DC_AO, response = "delta_14C", predictor = "Ca_mg_gSoil_DC")
```


```{r}
MAOM_14C_univariate <- HLEF_HFF_14C_DC_AO |> 
  dplyr::select(delta_14C, SOC_mgC_gSoil, TOC, ph, depth_cm, depth_from_org, 
                HPU, Fe_mg_gSoil_DC)
readr::write_csv(MAOM_14C_univariate, here("Data/dataframes/MAOM_14C_univariate.csv"))
```


